---
title: "Inference Statistics"
author: "Israel Almodovar-Rivera, PhD"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
  extract <- function(text) {
    text <- gsub(" ", "", text)
    split <- strsplit(text, ",", fixed = FALSE)[[1]]
    as.numeric(split)
  }

```

This *Shiny App* is made with the purpose to be free and educational. It comes without guarantee. You can perform basic statistical inference analysis such as one-sample and two-sample means comparison and comparison of variance. To learn more, see [FIRSTkit](https://github.com/ialmodovar/FIRSTkit).


```{r inputs, echo=FALSE}

sidebarLayout(
  sidebarPanel(
    tags$b("Data"),
      textInput("x", "Group 1", value = "", placeholder = "Enter values separated by a comma with decimals as points, e.g. 4.2, 4.4, 5, 5.03, etc."),
      textInput("alpha","alpha",value="0.05",placeholder="Significance level"),
      textInput("mu0","Mean",value="0",placeholder="Number indicating the true value of the mean "),
    checkboxInput("moregroups", "More groups?", FALSE),
      textInput("y", "Group 2", value = "", placeholder = "Enter values separated by a comma with decimals as points, e.g. 4.2, 4.4, 5, 5.03, etc."),
      hr()
    ),
  mainPanel(
    tabsetPanel(type = "tabs",
    tabPanel("Dataset", DT::dataTableOutput('tbldata')),
    tabPanel("One-Sample test", uiOutput("onesample")),
    tabPanel("Two-Sample test", uiOutput("twosample")),
    tabPanel("Three or more sample inference", uiOutput("ksample")))              )
)

## Data output

output$tbldata <- DT::renderDataTable({
    x <- as.numeric(extract(input$x))
    if(input$moregroups){
    y <- as.numeric(extract(input$y))
     DT::datatable(data.frame("Group 1" = x, "Group 2" = y))
    } else{
       DT::datatable(data.frame("Group 1" = x))
    }
  })
  
output$onesample <- renderUI({
    x <- as.numeric(extract(input$x))
    y <- as.numeric(extract(input$y))
    mu0 <- as.numeric(input$mu0)
    alpha <- as.numeric(input$alpha)
    z <- x
#    if(input$moregroups){
      twosided <- t.test(x = z, mu = mu0, 
                         alternative = "two.sided",
                         conf.level = 1-alpha)
      oneless <- t.test(x = z, mu = mu0, 
                         alternative = "less",
                         conf.level = 1-alpha)
      onegreater <- t.test(x = z, mu = mu0, 
                         alternative = "greater",
                         conf.level = 1-alpha)
   # }
})
  
```

## One Sample inference Hypothesis testing

### Mean

$H_0: \mu = \mu_0$ versus $H_1: \mu \neq \mu_0$

### Variance

$H_0: \sigma = \sigma_0$ versus $H_1: \sigma \neq \sigma_0$


## Two Sample inference Hypothesis testing

### Mean

$H_0: \mu_1 - \mu_2 = \delta$ versus $H_1: \mu \neq \mu_0$

### Variance

$H_0: \sigma = \sigma_0$ versus $H_1: \sigma \neq \sigma_0$


## Three or more Sample Inference Hypothesis testing

### Mean

$H_0: \mu_1 = \mu_2 = \ldots = \mu_k$ versus $H_1:$ at least one of the $\mu_i$ is different

### Variance

$H_0: \sigma = \sigma_0$ versus $H_1: \sigma \neq \sigma_0$


Report any bug to *israel.almodovar@upr.edu*
