---
title: "One-Sample Inference"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
library("readxl")
knitr::opts_chunk$set(echo = TRUE)
  extract <- function(text) {
    text <- gsub(" ", "", text)
    split <- strsplit(text, ",", fixed = FALSE)[[1]]
    as.numeric(split)
  }

all.t.one.sample.test <- function(x, mu0 = 0, alpha = 0.05,...){
  twosided <- t.test(x = x, mu = mu0, 
                     alternative = "two.sided",
                     conf.level = 1-alpha)
  oneless <- t.test(x = x, mu = mu0, 
                    alternative = "less",
                    conf.level = 1-alpha)
  onegreater <- t.test(x = x, mu = mu0, 
                       alternative = "greater",
                       conf.level = 1-alpha)
  
tt <- round(c(twosided$statistic, oneless$statistic, onegreater$statistic),digits=3)
pval <- round(c(twosided$p.value, oneless$p.value, onegreater$p.value),digits = 5) 

ci <- c(paste("(",paste(round(twosided$conf.int[1],digits=3),",",
        round(twosided$conf.int[2],digits=3),sep=""),")",sep=""),
        paste("(",paste(round(oneless$conf.int[1],digits=3),",",
                        round(oneless$conf.int[2],digits=3),sep=""),")",sep=""),
        paste("(",paste(round(onegreater$conf.int[1],digits=3),",",
                        round(onegreater$conf.int[2],digits=3),sep=""),")",sep=""))

res <- data.frame(Test = tt,Pval = pval, CI =ci)
names(res) <- c("statistic","p-value",paste(as.character((1-alpha)*100),"% CI",sep=""))
 rownames(res) <- c(paste("H_1: mu = ",mu0,sep=""),
                     paste("H_1: mu < ",mu0,sep=""),
                     paste("H_1: mu > ",mu0,sep=""))
res
}

all.wilcoxon.one.sample.test <- function(x, mu0 = 0, alpha = 0.05,...){
  twosided <- wilcox.test(x = x, mu = mu0, 
                     alternative = "two.sided",
                     conf.level = 1-alpha,conf.int = TRUE)
  oneless <- wilcox.test(x = x, mu = mu0, 
                         alternative = "less",
                         conf.level = 1-alpha,conf.int = TRUE)
  onegreater <- wilcox.test(x = x, mu = mu0, 
                            alternative = "greater",
                            conf.level = 1-alpha,conf.int = TRUE)
  
  tt <- round(c(twosided$statistic, oneless$statistic, onegreater$statistic),digits=3)
  pval <- round(c(twosided$p.value, oneless$p.value, onegreater$p.value),digits = 5) 
  
  ci <- c(paste("(",paste(round(twosided$conf.int[1],digits=3),",",
                          round(twosided$conf.int[2],digits=3),sep=""),")",sep=""),
          paste("(",paste(round(oneless$conf.int[1],digits=3),",",
                          round(oneless$conf.int[2],digits=3),sep=""),")",sep=""),
          paste("(",paste(round(onegreater$conf.int[1],digits=3),",",
                          round(onegreater$conf.int[2],digits=3),sep=""),")",sep=""))
  
  res <- data.frame(Test = tt,Pval = pval, CI =ci)
  names(res) <- c("statistic","p-value",paste(as.character((1-alpha)*100),"% CI",sep=""))
  
 rownames(res) <- c(paste("H_1: mu = ",mu0,sep=""),
                     paste("H_1: mu < ",mu0,sep=""),
                     paste("H_1: mu > ",mu0,sep=""))
 
res
}

one.sample.var.test <- function(x,sigma0=1,alpha=0.05,...){
  x <- x[!is.na(x)]
    df <- length(x)-1
  vv <- var(x,na.rm = TRUE)
  sigma02 <- sigma0^2
  X2 <- df * vv/sigma02 #test statistics
##
  pval <- c(pchisq(X2,df=df,lower.tail = FALSE),pchisq(X2,df=df,lower.tail = TRUE),pchisq(X2,df=df,lower.tail = FALSE))
  c1 <- qchisq(p=1-alpha/2,df = df)
  c2 <- qchisq(p=alpha/2,df = df)
  ci <- c(paste("(",round(df*vv/c1,digits=3),",",round(df*vv/c2,digits=3),")",sep=""),paste("(",0,",",round(df*vv/qchisq(p=alpha,df = df),digits=3),")",sep=""),
       paste("(",round(df*vv/qchisq(p=alpha,df = df),digits=3),",",Inf,")",sep=""))
  res <- data.frame(Test = X2,Pval = pval, CI =ci)
names(res) <- c("statistic","p-value",paste(as.character((1-alpha)*100),"% CI",sep=""))

 rownames(res) <- c(paste("H_1: sigma^2 =",sigma0,sep=""),
                     paste("H_1: sigma^2 < ",sigma0,sep=""),
                     paste("H_1: sigma^2 > ",sigma0,sep=""))
 
res

}

one.sample.prop.test <- function(x = NULL,n, p = NULL, p0 = 0.5, alpha = 0.05)
{
  if(is.null(x)& is.null(p)){
    stop("You must either provide the number of success or the probability of success")
  }
  if(!is.null(x) & is.null(p)){
    p <- x/n
  } 
  if(is.null(x) & !is.null(p)){
    p <- p
  }
if(!is.null(x) & !is.null(p)){
    p <- x/n
  } 
  z <- (p-p0)/sqrt(p0*(1-p0)/n)
  pval <- c(2 * pnorm(abs(z), lower.tail = FALSE),pnorm(z, lower.tail = TRUE),pnorm(z, lower.tail = FALSE))
  se <- sqrt(p*(1-p)/n)
  ci <- c(paste("(",paste(round(p+c(-1,1)* qnorm(1-alpha/2) * se,digits=5),collapse = ","),")",sep=""),
          paste("(",0,",",round(p-1* qnorm(1-alpha) * se,digits=5),")",sep=""),
paste("(",round(p+1* qnorm(alpha) * se,digits=5),",",1,")",sep=""))
  
  res <- data.frame(Test = z,Pval = pval, CI =ci)
names(res) <- c("statistic","p-value",paste(as.character((1-alpha)*100),"% CI",sep=""))

 rownames(res) <- c(paste("H_1: p = ",p0,sep=""),
                     paste("H_1: p < ",p0,sep=""),
                     paste("H_1: p > ",p0,sep=""))
 
res

}

```

This **Shiny App** is made with the purpose to be free and educational. It comes without guarantee. You can perform basic statistical inference analysis such as one-sample. The user can perform parametric *t* test as well, $\chi^2$ test for the variance. Note that the argument to perform a one sample test for the variance is in terms of the standard deviation. 

User can also select non-parametric approach. Wilcoxon-signed rank test uses the argument of $\mu_0$, although, this does not represent the mean but rather a location parameter. 


```{r inputs-one-sample, echo=FALSE}

sidebarLayout(
  sidebarPanel(
    selectInput(
      inputId = "datachoice",
      label = "Data (Input some data or upload a file)",
      choices = c("Input data","Upload file","Upload .xls/.xslx file"),
      multiple = FALSE,
      selected = NULL##"Input data"
    ),
    hr(),
    conditionalPanel(
    condition = "input.datachoice == 'Input data'",
    textInput("x", "Data", value = "",
    placeholder = "Enter values separated by a comma with decimals as points, e.g. 1.2, 3.14, 3.1417, etc."),
    hr(),
    ),
    conditionalPanel(
    condition = "input.datachoice == 'Upload file'",
    fileInput("file1", "Choose CSV File",
              multiple = TRUE,
              accept = c("text/csv",
                         "text/comma-separated-values,text/plain",
                         ".csv")),
    hr(),
        # Input: Checkbox if file has header ----
    checkboxInput("header", "Header", TRUE),
        
    # Input: Select separator ----
    radioButtons("sep", "Separator",
                  choices = c(Comma = ",",
                              Semicolon = ";",
                              Tab = "\t"),
                  selected = ","),
        
        # Input: Select quotes ----
    radioButtons("quote", "Quote",
                 choices = c(None = "",
                            "Double Quote" = '"',
                            "Single Quote" = "'"),
                selected = '"'),
    ),
    conditionalPanel(
    condition = "input.datachoice == 'Upload .xls/.xslx file'",
    fileInput('file2', 'Choose .xls/.xlsx file',accept = c(".xlsx",".xls"))
    ),
    selectInput("variable","Variable", choices = NULL),
    uiOutput("alpha_line"),
    textInput("mu0","\\(  H_0:\\mu = \\mu_0 \\)",value="0",placeholder="Indicate the true value for the location of interest "),
     textInput("sigma0","\\( H_0:\\sigma^2 = \\sigma^2_0 \\)",value="1",placeholder="Indicate the true value for the dispersion of interest"),
    textInput("p0","\\( H_0: p = p_0 \\)",value="0.5",placeholder="Indicate the true value for the proportion of interest"),
        numericInput("nprop", " Sample size (\\(n \\))",
                     value = 50, min = 0, step = 1
        ),
            hr(),
        radioButtons(
          inputId = "propsuccess",
          label = NULL,
          choices = c(
            "Proportion of success \\(\\hat{p}\\)" = "pr",
            "Number of successes \\(x\\)" = "success"
          )
        ),
       conditionalPanel(
          condition = "input.propsuccess == 'pr'",
          numericInput("p_prop", "Proportion of success (\\(\\hat{p} \\))",
                       value = 0.5, min = 0, max = 1, step = 0.001
          )
        ),
        conditionalPanel(
          condition = "input.propsuccess == 'success'",
          numericInput("x_prop", "Number of successes(\\(x\\))",
                       value = 20, min = 0, step = 1
          )
        ),
        sliderInput("alpha","Significance level \\(\\alpha \\)",
                    min = 0,
                    max = 1,
                    value = 0.05,step=0.001,
        ),

    ),
      mainPanel(
        tabsetPanel(type = "tabs",
        tabPanel("Location Inference",
        br(),
        tags$b("One Sample \\( t \\) test"),
        uiOutput("mean"),
        br(),
        tags$b("Wilcoxon-signed rank test"),
        uiOutput("loc")),
        tabPanel("Dispersion Inference",
        br(),
        tags$b("One Sample Variance test"),
        withMathJax(uiOutput("variance"))),
        tabPanel("Proportion Inference",
        tags$b("One Sample proportion test"),
                 uiOutput("proportion"))
        )
  )
    )

d <- reactive({  
  if(!is.null(input$file1)){
    req(input$file1)
    
    tt <- read.csv(input$file1$datapath,
                   header = input$header,
                   sep = input$sep,
                   quote = input$quote)
   
  } 
  if(!is.null(input$file2)){
    req(input$file2)
    tt <- as.data.frame(read_excel(input$file2$datapath,sheet=1))
  }
  if(is.null(input$file1) & is.null(input$file2)){
  x <- extract(input$x)
  tt<- data.frame(x = x)
  } 
  
  var.opts <- colnames(tt)
  updateSelectInput(inputId =  "variable", choices = var.opts)

  return(tt)
})


output$mean <- renderTable({
    mu0 <- as.numeric(input$mu0)
    alpha <- as.numeric(input$alpha)
    df <- d()
    
    x <- df[,input$variable]
    
   all.t.one.sample.test(x = x, mu0 = mu0,alpha = alpha)
},rownames = TRUE)

output$loc <- renderTable({
    mu0 <- as.numeric(input$mu0)
    alpha <- as.numeric(input$alpha)
    df <- d()
    
    x <- df[,input$variable]
    
   all.wilcoxon.one.sample.test(x = x, mu0 = mu0,alpha = alpha)
   
},rownames = TRUE)

output$variance <- renderTable({
    sigma0 <- as.numeric(input$sigma0)
    alpha <- as.numeric(input$alpha)
    df <- d()
    
    x <- df[,input$variable]
    
   one.sample.var.test(x = x, sigma0 = sigma0,alpha = alpha)
},rownames = TRUE)

output$proportion <- renderTable({
  p0 <- as.numeric(input$p0)
  alpha <- as.numeric(input$alpha)
  
  one.sample.prop.test(x = input$x_prop,
                       n = input$nprop,p = input$p_prop,p0 = p0,alpha = alpha)
},rownames = TRUE,digits = 4)
  
```

### One sample $t$ test

### A chi-square test

A chi-square test (Snedecor and Cochran, 1983) can be used to test if the variance of a population is equal to a specified value. This test can be either a two-sided test or a one-sided test. The two-sided version tests against the alternative that the true variance is either less than or greater than the specified value. The one-sided version only tests in one direction. The choice of a two-sided or one-sided test is determined by the problem. For example, if we are testing a new process, we may only be concerned if its variability is greater than the variability of the current process. 

To learn more, see [FIRSTkit](https://github.com/ialmodovar/FIRSTkit). If you have any question or want to report to *israel.almodovar@upr.edu* or *maitra@iastate.edu*. 
